---
title: "Barcode"
output: html_document
date: "2022-12-20"
---

```{r}
getwd()
library("tidyverse")

```

miRNAseed sequence is in a .fa, so use Winzip to open and then convert to .txt format Next issue is that the format uses two rows per information, which doesn't work for reading into a table

Solution:

```{r}
link <-"~/allmiRNAseed.txt"
##Read the raw lines
rawlines <-readLines(file(link))
##Convert to one text string
lines <- paste(rawlines[c(TRUE, FALSE)], rawlines[c(FALSE, TRUE)], 
               collapse = "\n")
##paste is used to combine multiple 
##rawlines[c(TRUE, FALSE)] represents the odd lines, 
##rawlines[c(FALSE, TRUE)] represents the even lines.
##Both small lines are combined to one long line. 
##Then, all long lines are combined into one single string 
##with the argument collapse = "\n". 
##The lines are separated by line breaks (\n).
dat <- read.table(text = lines, na.string = "*") %>%
unite(specie, V3, V4) 
dat1 <-filter(dat, specie == "Homo_sapiens")
print(dat1)
```

```{r}
setwd("C:/Users/proli/OneDrive - The Pennsylvania State University/2022 Carrel Liu Lab/Alzheimers Computational/2022 Barcode")
library("Biostrings")
library("tidyverse")
library("immunarch")
library("universalmotif")
nucruns <- vector(mode = 'character', length = 4) %>% DNAStringSet
ni = 1
for (i in 4) {
  for (j in c('A', 'G', 'T', 'C')) {
    nucruns[ni] = rep(j, i) %>% paste(collapse = '') %>% DNAStringSet
    ni = ni + 1
  }
}

##not sure what we have done
twelvemers <- create_sequences("DNA", 200, 10) %>% DNAStringSet 
cat(paste0('done generating 12mers at ', Sys.time()))

#Each nucleotide occurs at least one
missingone <- apply(alphabetFrequency(twelvemers)[,1:4], 1, function(x){any(x == 0)})
twelvemers <- twelvemers[!missingone]
cat(paste0('done removing twelvemers missing a nucleotide at ', Sys.time()))

#Cut out those with nucleotide runs of 4 or more in a row
hasnucruns <- vcountPDict(nucruns, twelvemers) %>% colSums
hasnucruns <- hasnucruns > 0
twelvemers <- twelvemers[!hasnucruns]
cat(paste0('done removing 12mers with runs of 4 or more at ', Sys.time()))
print(twelvemers)

#Cut out those that match the miRNA seed sequences 
#For now let's just use the human ones since there are fewer and it won't take as long
#isolated species names with cat mature.fa | grep '>' | cut -f 3,4 -d \  > mirBaseSpecies.txt

species <- dat1 %>% 
   tibble::as_tibble() %>% 
   transmute(name = paste(V1 %>% as.character, V2 %>% as.character)) 

human = grepl('Homo sapiens', species$name)
allSeeds <- readRNAStringSet("~/mature.fa")
##download mature.fa from: https://www.mirbase.org/ftp.shtml
seedSeqs <- allSeeds %>% subseq(2,7) %>% DNAStringSet %>% unique
humanSeedSeqs <- seedSeqs[human] %>% unique
seedSeqs <- allSeeds %>% subseq(2,7) %>% DNAStringSet %>% unique

haveSeedlist <- vwhichPDict(humanSeedSeqs, twelvemers) 
#this takes ~40 minutes. All seeds takes ~1h45m
save(list = c('twelvemers', 'haveSeedlist', 'humanSeedSeqs'), 
     file = '~/haveHumanRNAiSeeds.RData')
haveSeed <- sapply(haveSeedlist, function(x){length(x) > 0})
twelvemers <- twelvemers[!haveSeed]
cat(paste0('done removing those with mirSeeds at ', Sys.time()))
print(length(twelvemers))


```

Sandbox

```{r}
setwd("C:/Users/proli/OneDrive - The Pennsylvania State University/2022 Carrel Liu Lab/Alzheimers Computational/2022 Barcode")
library("Biostrings")
library("tidyverse")
library("immunarch")
library("universalmotif")
nucruns <- vector(mode = 'character', length = 4) %>% DNAStringSet
ni = 1
for (i in 4) {
  for (j in c('A', 'G', 'T', 'C')) {
    nucruns[ni] = rep(j, i) %>% 
      paste(collapse = '') %>% 
      DNAStringSet
    ni = ni + 1
  }
}


twelvemers <- create_sequences("DNA", 200, 10, rng.seed = sample.int(10000, 1)) %>% 
  DNAStringSet 
cat(paste0('done generating 12mers at ', Sys.time()))

#Each nucleotide occurs at least once
missingone <- apply(alphabetFrequency(twelvemers)[,1:4], 1, function(x){any(x == 0)})
twelvemers <- twelvemers[!missingone]
cat(paste0('done removing twelvemers missing a nucleotide at ', Sys.time()))

#Cut out those with nucleotide runs of 4 or more in a row
hasnucruns <- vcountPDict(nucruns, twelvemers) %>% 
  colSums
hasnucruns <- hasnucruns > 0
twelvemers <- twelvemers[!hasnucruns]
cat(paste0('done removing 12mers with runs of 4 or more at ', Sys.time()))
print(twelvemers
      )

#Cut out those that match the miRNA seed sequences 
#For now let's just use the human ones since there are fewer and it won't take as long
#isolated species names with cat mature.fa | grep '>' | cut -f 3,4 -d \  > mirBaseSpecies.txt

species <- dat1 %>% 
   tibble::as_tibble()
print(species)
#convert dat1 to tibble named "species"

human <- grepl('Homo sapiens', species$specie)
#looks for Home sapiens in species file and pulls those, redundant
print(human)
allSeeds <- readRNAStringSet("~/mature.fa")
#download mature.fa from: https://www.mirbase.org/ftp.shtml
seedSeqs <- allSeeds %>% subseq(2,7) %>% DNAStringSet %>% unique
humanSeedSeqs <- seedSeqs[human] %>% unique
seedSeqs <- allSeeds %>% subseq(2,7) %>% DNAStringSet %>% unique

haveSeedlist <- vwhichPDict(humanSeedSeqs, twelvemers) 
#this takes ~40 minutes. All seeds takes ~1h45m
save(list = c('twelvemers', 'haveSeedlist', 'humanSeedSeqs'), 
     file = '~/haveHumanRNAiSeeds.csv')
haveSeed <- sapply(haveSeedlist, function(x){length(x) > 0})
twelvemers <- twelvemers[!haveSeed]
cat(paste0('done removing those with mirSeeds at ', Sys.time()))
print(length(twelvemers))

save(twelvemers, file = '~/inertTwelveMers.csv')
